Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-04-07 demystify.tex
typos

\textbf{Advice}: do not spend much time talking about non-research-related job in your SOP.
However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs. For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potentials in their LOR (\autoref{sec:lor}).

@@

\paragraph{Interviews} Just like a job application, you likely will get interviews from schools. The most common case is that a prof. is interested in working with you and wants to chat to make decision, e.g., to offer RA (\autoref{sec:ra})). In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program. The interview can also be done by adcom members to see if the student fit their program.
%The professor might include one of their current students as part of the interview.

Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
You might be asked to talk about your research experiences and interests, and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
The interview is also a good opportunity for you to ask questions about the program and the university.

\textbf{Advice}: Treat the interview as an informal chat. Have elevator pitches about your research experiences and interests. You might also want have some 5-minute presentation about your research. If a prof. asks you to read a paper, do it and be prepared to discuss it. You can also ask if you need to prepare for a coding problem. Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research.

\begin{commentbox}[Vu:]
  At GMU, faculty are encouraged to interview candidates. For very strong candidates, the interview is actually to recruit them.  In some cases a faculty interviews a candidate that they see potentials and want to advocate for their admission. Without the interview, such applications may be more likely to be rejected.
  \tcblower
  In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
\end{commentbox}

@@

\paragraph{Not getting interviews} While it is generally good to get interview, not getting an interview \emph{does not} mean you're out.  Many programs do not have the tradition of interviewing applicants. For example, at GMU, most of our admitted students with TA (\autoref{sec:ta}) do not go through interviews.

@@

\paragraph{Low Stipend?} Students often complain their stipend being too low. However, in most cases it is actually not bad and you can live comfortably with it.  It might also be enough to support your spouse and kids (many CS PhD students have their family with them). So don't worry too much about stipend.  If you're admitted to a good CS PhD program, you will be fine. A good school would know that it has to be competitive to attract students.  For example, at GMU, every year we discuss about improving the benefits, and especially stipend, for our graduate students.

@@

\section{Why did I get rejected?}\label{sec:why-rejected}
Many possible reasons!  Other than not having a strong profile (e.g., research potentials, GPA, LoRs, SOP, poor interviews), here are some other reasons.

>>>>

\textbf{Advice}: do not spend much time talking about non-research-related job in your SOP.
However, if you have research-based work experiences, then they can be useful and should be mentioned in your SOP and LORs. For example, if you have worked at a research lab, e.g., Microsoft, DoE, VinAI in Vietnam, etc (\autoref{sec:research-experiences}), then you should mention that in your SOP and ask your supervisor to write about your research experiences and potential in their LOR (\autoref{sec:lor}).

@@

\paragraph{Interviews} Just like a job application, you likely will get interviews from schools. The most common case is that a prof. is interested in working with you and wants to chat to make a decision, e.g., to offer RA (\autoref{sec:ra})). In some cases, the interview is done by several professors, e.g., to see if a student fits in their group or to recruit a very strong student to their program. The interview can also be done by adcom members to see if the student fits their program.
%The professor might include one of their current students as part of the interview.

Typically, an interview takes about 15--30 minutes, and one important aspect of evaluation is your ability to effectively communicate, including speaking and understanding English. 
You might be asked to talk about your research experiences and interests and to read a paper and discuss it. In some rare cases you might also be asked to solve a problem, e.g., a coding problem. 
The interview is also a good opportunity for you to ask questions about the program and the university.

\textbf{Advice}: Treat the interview as an informal chat. Have elevator pitches about your research experiences and interests. You might also want to have a 5-minute presentation about your research. If a prof. asks you to read a paper, do it and be prepared to discuss it. You can also ask if you need to prepare for a coding problem. Finally, prepare some thoughtful questions to ask, e.g., about the program, the university, or the professor's research.

\begin{commentbox}[Vu:]
  At GMU, faculty are encouraged to interview candidates. For very strong candidates, the interview is actually to recruit them.  In some cases a faculty interviews a candidate that they see potential and want to advocate for their admission. Without the interview, such applications may be more likely to be rejected.
  \tcblower
  In short, getting an interview is a good sign; it means that someone is considering you. If we are not interested in your application, we will not proceed with an interview.
\end{commentbox}

@@

\paragraph{Not getting interviews} While it is generally good to get an interview, not getting one \emph{does not} mean you're out.  Many programs do not have the tradition of interviewing applicants. For example, at GMU, most of our admitted students with TA (\autoref{sec:ta}) do not go through interviews.

@@

\paragraph{Low Stipend?} Students often complain their stipend being too low. However, in most cases it is actually not bad, and you can live comfortably with it.  It might also be enough to support your spouse and kids (many CS PhD students have their family with them). So don't worry too much about stipend.  If you're admitted to a good CS PhD program, you will be fine. A good school would know that it has to be competitive to attract students.  For example, at GMU, every year we discuss improving the benefits, and especially stipend, for our graduate students.

@@

\section{Why did I get rejected?}\label{sec:why-rejected}
Many possible reasons!  Other than not having a strong profile (e.g., research potential, GPA, LoRs, SOP, poor interviews), here are some other reasons.
